Improvement of crash forces in structures using optimization tools.

### 2016-07-28 paper.tex
Avances

\begin{abstract}
This work describes an investigation on the structural optimization of the crash response of aircraft or road vehicle crash boxes, where the force pulse due to frontal impact or hard landing wanted to be kept as stable as possible while reducing the mass of the energy absorbing systems. The objective functions are minimized in a multi-objective approach using metamodeling on FE models and genetic evolutionary algorithms. Finite element models were subjected to loading conditions simulating, for example, an impact test against a simplified rigid wall in the case of the car model, or a hardlanding impact for the fuselage model. Force-displacement curves were obtained from the analysis and suitably filtered. The objective was to obtain a force-time curve which was as close as possible to the ideal dissipator curve, with lower values of the peaks of acceleration experienced by the occupants; and to minimize the mass for fuel saving and environmental reasons. To that end, optimization strategies were planned carefully to deal with problems which are typical in crashworthiness optimization like expensive computation times and numerical noise.
\end{abstract}\\
\emph{Keywords: crashworthiness, injuries reduction, size optimization, surrogate models.}

@@

\begin{figure}[htpb]
\begin{subfigure}[b]{.45\columnwidth}
   \includegraphics[width=\columnwidth]{figures/IMG/ace}
    \caption{ACE body structure design by Honda. Relevant structural members and load paths. Taken and modified from \cite{tesis_mcostas}.}
    \label{fig:car}
\end{subfigure}
\qquad
\begin{subfigure}[b]{.45\columnwidth}
   \includegraphics[width=\columnwidth]{figures/IMG/energy_absorption2}
    \caption{Schematics of an aircraft fuselage drop test for crashworthiness evaluation. Taken from \cite{Paz:2014dw}.}
    \label{fig:plane}
\end{subfigure}
\label{fig:car_plane}
\end{figure}

% As shown if \cref{fig:car_plane},

\Cref{fig:car,fig:plane} show the schematics of a car frame and a fuselage section, both enclosing crash boxes. For these applications, adhesives are very competitive both in bonding strength and in economical manufacturing costs \cite{superyacht}. In order to check adhesive's suitability for their use in crashworthy elements, this research develops numerical models of adhesively bonded crash boxes subjected to impact loads using finite element models. Information and data present in the literature are used to create and validate these models in order to ensure the accuracy of the results.

\section{Finite element model}
The modeled crash box consists of two adherend cold-formed sheets made of steel that create a tube of square section bonded together with an epoxy adhesive. The joint is practiced on the surface of two flanges on each adherend side. In order to ease the model validation, the crash box geometry was created to reflect those tested by \cite{Peroni2009} and later modeled by \cite{Scattina2011}. In particular, the two sections depicted in \cref{fig:crash_box} are modeled in the present study.
\begin{figure}[ht]
  \centering
  \begin{minipage}[b]{.48\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/IMG_CUTRES/medidas_cb}
    \subcaption{Dimensions of the modeled crash boxes, in millimeters. Taken from \cite{Peroni2009}.}
    \label{fig:crash_box}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{.48\linewidth}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/IMG_CUTRES/general_transp}
    \subcaption{General view of the model with the bounding box.}
    \label{fig:general}
  \end{minipage}
  \caption{Dimensions of the studied crash boxes and view of the finite element model including a bounding box.}
  \label{fig:modelo}
\end{figure}

@@

The last $\SI{5}{\mm}$ of the crash tube had all degrees of freedom fixed (see \cref{fig:general}), except displacement on the impact direction in order to allow the reaction force measurement commented before. This way, numerical issues due to excesive unrestrained degrees of freedom on the whole tube were prevented.

Peeling problems were found near the impact head end during the development of the study. \cite{Peroni2009} solved this situation by adding two rivets on the bonded flanges at the triggered section in order to avoid excesive debonding on that part of the tube in the initial phases of the impact and, thus, prevent critical situations. In this study, this was simulated by a $\SI{2}{\mm}$-diameter spot-weld at the same location.

@@

According to \cite{manufCatalog}, the shear strength of the adhesive is $\SI{44}{\MPa}$, although this parameter is subject to variation depending both in cure temperature and time. The bulk modulus is $\SI{1460}{\MPa}$ and the adhesive's density is $\SI{1.30}{\tonne/\m^3}$ \cite{manufCatalog}. The constitutive behavior was supposed to be isotropic linear elastic \cite{SernaMoreno2015} up to failure start. An uncoupled traction elastic behavior \cite{Sadowski2010, Sadowski2011, Scattina2011, Sadowski2014} on 3D cohesive finite elements was chosen.

\cite{Scattina2011} obtained adhesive stiffnesses through the inverse method, which are summarized in \cref{tab:ads_params}. These same parameters were used in the present study, as their results could be validated with the model.
\begin{table}
	\centering
	\begin{tabular}{llrl}
	\hline
		Parameter & Description & Value & \\
		\hline
		$E_{N}$ & Stiffness in normal direction & $\num{5e9}$ & $\si{\kN/\m^2}$ \\
		$E_{T}$ & Stiffness in an in-plane direction & $\num{8e7}$ & $\si{\kN/\m^2}$ \\
		$G_{Ic}$ & Energy release rate for mode I & $\num{2028}$ & $\si{\J/\m^2}$ \\
		$G_{IIc}$ & Energy release rate for mode II & $\num{11853}$ & $\si{\J/\m^2}$ \\
		$\sigma_{n}^{0}$ & Peak traction in normal direction & $\num{130}$ & $\si{\MPa}$ \\
		$\tau_{II}^{0}$ & Peak traction in tangential direction & $\num{42.5}$ & $\si{\MPa}$ \\
		\hline
	\end{tabular}
	\caption[Loctite Hysol 9514 parameters.]{Loctite Hysol 9514 parameters. Taken from \cite{Scattina2011}.}
	\label{tab:ads_params}
\end{table}

Contact failure modeling ---corresponding to adhesive failure--- resulted in some simulation problems, so adhesives and adherends were kept together with a simplified tie constraint. Thus, failure definitions were only included in the bulk material, obviating the contact/adhesive failure as suggested by several authors \cite{Greve2007, Loureiro2010, Sadowski2010, Sadowski2011, Scattina2011, Sadowski2014, SernaMoreno2015}, although the modeled cohesive damage parameters include the effect of both adhesive and cohesive resistance.

@@

% Tie (include image)
The tie constraint included in Abaqus enabled non-coherent meshes between adhesive and adherend (\cref{fig:mesh_detail_coh3d_comparison}). This way, finer meshes could be implemented on the adhesive in order to improve the representation of the failure progression without increasing the adherend element recount by leaving a coarser mesh there.
\begin{figure}[htpb]
	\centering
	\begin{minipage}[b]{.7\linewidth}
		\centering
		\includegraphics[width=0.7\linewidth]{figures/IMG_CUTRES/mesh_detail_coh3d_comparison}
		\caption[Detail of COH3D mesh on the adhesive layer.]{Detail of COH3D mesh on the adhesive layer (blue).}
		\label{fig:mesh_detail_coh3d_comparison}
	\end{minipage}
	\\[10pt]
	\begin{minipage}[b]{.7\linewidth}
		\centering
		\includegraphics[width=\linewidth]{figures/IMG_CUTRES/mesh}
		\caption{General mesh of the crash box.}
		\label{fig:mesh}
	\end{minipage}
	\caption{Views of the mesh of the finite element model: globral view and detail view of the adhesive.}
	\label{fig:mallas}
\end{figure}

The finite element package Abaqus Explicit is used in this research, on is version 6.13 \cite{Abaqus613Manual} running on a high performance computing cluster. Jobs running on loop parallelization over $16$ cores took about $\SI{8}{\hour}$ to finish. Domain parallelization turned out to be inefficient in this case due to poor network performance.

@@

In order to filter high frequency effects, a SAE 600 filter was applied to the simulation results \cite{Huang}. \Cref{fig:F-D_rough} provides the obtained force-displacement curves and the results obtained in \cite{Peroni2009}. In this figure the nomenclature is as follows:
\begin{itemize}
	\item ``AR'' or ``BR'' indicating the cross-section of the tube: ``A'' for the top-hat, and ``B'' for the double hat.
	\item An ``R'' after the underline indicates that the crash box had riveting in the triggered cross section.
	\item ``Sb'' after the underline marks the presence of a stabilizing box in the model.
\end{itemize}

@@

Once the sampling is obtained, the surrogate model is created. The Multivariate adaptive regression splines (MARS) method is chosen, where the surrogate functions are adjusted with cubic splines following the next expression
\begin{equation}\label{eq:mars}
\hat{f}\left ( \bm{x} \right )= \sum_{m=1}^{M}a_{m}B_{m}\left ( \bm{x} \right ),
\end{equation}
where $B_{m}$ are the power basis functions, $a_{m}$ the coefficients of the functions, and $M$ is the number of functions. This method is more thoroughly explained in \cite{Friedman1995197}.

@@

Furthermore, other researchers have traditionally optimized the specific energy absorption (SEA) and the load ratio (LR) \cite{Hou2007555}. Thus, they are also evaluated as objective functions, but considering their more complicated comportment and a noisier nature for this model.

>>>>

\begin{abstract}
This work describes an investigation on the structural optimization of the crash response of aircraft or road vehicle crash boxes, where the force pulse due to frontal impact or hard landing wanted to be kept as stable as possible while reducing the mass of the energy absorbing systems. The objective functions are minimized in a multi-objective approach using metamodeling on FE models and genetic evolutionary algorithms. Finite element models were subjected to loading conditions simulating, for example, an impact test against a simplified rigid wall in the case of the car model, or a hardlanding impact for the fuselage model. Force-displacement curves were obtained from the analysis and suitably filtered. The objective was to obtain a force-time curve which was as close as possible to the ideal dissipator curve, with lower values of the peaks of acceleration experienced by the occupants; and to minimize the mass for fuel saving and environmental reasons. To that end, optimization strategies were planned carefully to deal with problems which are typical in crashworthiness optimization like expensive computation times and numerical noise.
\end{abstract}\\
\emph{Keywords: crashworthiness, damage reduction, size optimization, surrogate models.}

@@

\begin{figure}[htpb]
\begin{subfigure}[b]{.45\columnwidth}
   \includegraphics[width=\columnwidth]{figures/IMG/ace}
    \caption{ACE body structure design by Honda. Relevant structural members and load paths. Taken from (Costas, \cite{tesis_mcostas}).}
    \label{fig:car}
\end{subfigure}
\qquad
\begin{subfigure}[b]{.45\columnwidth}
   \includegraphics[width=\columnwidth]{figures/IMG/energy_absorption2}
    \caption{Schematics of an aircraft fuselage drop test for crashworthiness evaluation. Taken from (Paz, \cite{Paz:2014dw}).}
    \label{fig:plane}
\end{subfigure}
\label{fig:car_plane}
\end{figure}

\Cref{fig:car,fig:plane} show the schematics of a car frame and a fuselage section, both enclosing crash boxes. For these applications, adhesives are very competitive both in bonding strength and in economical manufacturing costs (Jones, \cite{superyacht}). In order to check adhesive's suitability for their use in crashworthy elements, this research develops numerical models of adhesively bonded crash boxes subjected to impact loads using finite element models. Information and data present in the literature are used to create and validate these models in order to ensure the accuracy of the results.

\section{Finite element model}
The modeled crash box consists of two adherend cold-formed sheets made of steel that create a tube of square section bonded together with an epoxy adhesive. The joint is practiced on the surface of two flanges on each adherend side. In order to ease the model validation, the crash box geometry was created to reflect those tested by (Peroni, \cite{Peroni2009}) and later modeled by (Scattina, \cite{Scattina2011}). In particular, the two sections depicted in \Cref{fig:crash_box} are modeled in the present study.
\begin{figure}[ht]
  \centering
  \begin{minipage}[b]{.48\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/IMG_CUTRES/medidas_cb}
    \subcaption{Dimensions of the modeled crash boxes, in millimeters. Taken from \cite{Peroni2009}.}
    \label{fig:crash_box}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{.48\linewidth}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/IMG_CUTRES/general_transp}
    \subcaption{General view of the model with the bounding box.}
    \label{fig:general}
  \end{minipage}
  \caption{Dimensions of the studied crash boxes and view of the finite element model including a bounding box.}
  \label{fig:modelo}
\end{figure}

@@

The last $\SI{5}{\mm}$ of the crash tube had all degrees of freedom fixed (see \Cref{fig:general}), except displacement on the impact direction in order to allow the reaction force measurement commented before. This way, numerical issues due to excesive unrestrained degrees of freedom on the whole tube were prevented.

Peeling problems were found near the impact head end during the development of the study. This was solved in \cite{Peroni2009} by adding two rivets on the bonded flanges at the triggered section in order to avoid excessive debonding on that part of the tube in the initial phases of the impact and, thus, prevent critical situations. In this study, this was simulated by a $\SI{2}{\mm}$-diameter spot-weld at the same location.

@@

According to (Loctite, \cite{manufCatalog}), the shear strength of the adhesive is $\SI{44}{\MPa}$, although this parameter is subject to variation depending both in cure temperature and time. The bulk modulus is $\SI{1460}{\MPa}$ and the adhesive's density is $\SI{1.30}{\tonne/\m^3}$ \cite{manufCatalog}. The constitutive behavior was supposed to be isotropic linear elastic \cite{SernaMoreno2015} up to failure start. An uncoupled traction elastic behavior (Sadowski, Scattina \cite{Sadowski2010, Sadowski2011, Scattina2011, Sadowski2014}) on 3D cohesive finite elements was chosen.

\cite{Scattina2011} obtained adhesive stiffnesses through the inverse method, which are summarized in \Cref{tab:ads_params}. These same parameters were used in the present study, as their results could be validated with the model.
\begin{table}
	\centering
	\begin{tabular}{llrl}
	\hline
		Parameter & Description & Value & \\
		\hline
		$E_{N}$ & Stiffness in normal direction & $\num{5e9}$ & $\si{\kN/\m^2}$ \\
		$E_{T}$ & Stiffness in an in-plane direction & $\num{8e7}$ & $\si{\kN/\m^2}$ \\
		$G_{Ic}$ & Energy release rate for mode I & $\num{2028}$ & $\si{\J/\m^2}$ \\
		$G_{IIc}$ & Energy release rate for mode II & $\num{11853}$ & $\si{\J/\m^2}$ \\
		$\sigma_{n}^{0}$ & Peak traction in normal direction & $\num{130}$ & $\si{\MPa}$ \\
		$\tau_{II}^{0}$ & Peak traction in tangential direction & $\num{42.5}$ & $\si{\MPa}$ \\
		\hline
	\end{tabular}
	\caption[Loctite Hysol 9514 parameters.]{Loctite Hysol 9514 parameters. Taken from \cite{Scattina2011}.}
	\label{tab:ads_params}
\end{table}

Contact failure modeling ---corresponding to adhesive failure--- resulted in some simulation problems, so adhesives and adherends were kept together with a simplified tie constraint. Thus, failure definitions were only included in the bulk material, obviating the contact/adhesive failure as suggested by several authors (Greve, Loureiro, Sadowski, Scattina, Serna \cite{Greve2007, Loureiro2010, Sadowski2010, Sadowski2011, Scattina2011, Sadowski2014, SernaMoreno2015}), although the modeled cohesive damage parameters include the effect of both adhesive and cohesive resistance.

@@

% Tie (include image)
The tie constraint included in Abaqus enabled non-coherent meshes between adhesive and adherend (\Cref{fig:mesh_detail_coh3d_comparison}). This way, finer meshes could be implemented on the adhesive in order to improve the representation of the failure progression without increasing the adherend element recount by leaving a coarser mesh there.
\begin{figure}[htpb]
	\centering
	\begin{minipage}[b]{.7\linewidth}
		\centering
		\includegraphics[width=0.7\linewidth]{figures/IMG_CUTRES/mesh_detail_coh3d_comparison}
		\caption[Detail of COH3D mesh on the adhesive layer.]{Detail of COH3D mesh on the adhesive layer (blue).}
		\label{fig:mesh_detail_coh3d_comparison}
	\end{minipage}
	\\[10pt]
	\begin{minipage}[b]{.7\linewidth}
		\centering
		\includegraphics[width=\linewidth]{figures/IMG_CUTRES/mesh}
		\caption{General mesh of the crash box.}
		\label{fig:mesh}
	\end{minipage}
	\caption{Views of the mesh of the finite element model: globral view and detail view of the adhesive.}
	\label{fig:mallas}
\end{figure}

The finite element package Abaqus Explicit is used in this research, on is version 6.13 (Dassault, \cite{Abaqus613Manual}) running on a high performance computing cluster. Jobs running on loop parallelization over $16$ cores took about $\SI{8}{\hour}$ to finish. Domain parallelization turned out to be inefficient in this case due to poor network performance.

@@

In order to filter high frequency effects, a SAE 600 filter was applied to the simulation results (Huang, \cite{Huang}). \Cref{fig:F-D_rough} provides the obtained force-displacement curves and the results obtained in \cite{Peroni2009}. In this figure the nomenclature is as follows:
\begin{itemize}
	\item ``AR'' or ``BR'' indicating the cross-section of the tube: ``A'' for the top-hat, and ``B'' for the double hat.
	\item An ``R'' after the underline indicates that the crash box had riveting in the triggered cross section.
	\item ``Sb'' after the underline marks the presence of a stabilizing box in the model.
\end{itemize}

@@

Once the sampling is obtained, the surrogate model is created. The Multivariate adaptive regression splines (MARS) method is chosen, where the surrogate functions are adjusted with cubic splines following the next expression
\begin{equation}\label{eq:mars}
\hat{f}\left ( \bm{x} \right )= \sum_{m=1}^{M}a_{m}B_{m}\left ( \bm{x} \right ),
\end{equation}
where $B_{m}$ are the power basis functions, $a_{m}$ the coefficients of the functions, and $M$ is the number of functions. This method is more thoroughly explained in (Friedman, \cite{Friedman1995197}).

@@

Furthermore, other researchers have traditionally optimized the specific energy absorption (SEA) and the load ratio (LR) (Hou, \cite{Hou2007555}). Thus, they are also evaluated as objective functions, but considering their more complicated comportment and a noisier nature for this model.
